---
title: "Repetitive negative thinking following exposure to a natural stressor prospectively predicts altered stress reactivity and decision making in the laboratory"
output: 
  html_document:
    df_print: paged
    theme: cerulean
editor_options: 
  chunk_output_type: inline
---

####Caitlin A. Stamatis, Nikki Puccetti, Caroline Charpentier, Aaron S. Heller, & Kiara R. Timpano####

<br/>
This document contains the code for analyses conducted for a manuscript in submission.
<br/>

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo = TRUE)
library(stargazer)
library(sjPlot)
library(sjmisc)
library(ggplot2)
library(lmerTest)
library(lme4)
library(reshape2)
library(tidyr)
library(doBy)
library(psych)
library(dplyr)
library(nlme)
library(readxl)
library(summarytools)
library(knitr)
library(pander)
library(DT)
library(shiny)
library(plyr)
options(scipen = 999)
df <- read.csv("rntDecisionMakingData.csv")
df2 <- df
df2$Gender <- as.factor(df2$Gender)
df2$CurrentMed_notBC_t3 <- as.factor(df2$CurrentMed_notBC_t3)
df2$CurrentMed_BC_t3 <- as.factor(df2$CurrentMed_BC_t3)
df2$Smoker_0no_1yes_2yestoday_t3 <- as.factor(df2$Smoker_0no_1yes_2yestoday_t3)
df2$TodayCaff_NumCups_t3 <- as.factor(df2$TodayCaff_NumCups_t3)
df2$CurrentMenstr_t3 <- as.factor(df2$CurrentMenstr_t3)
```

###Data cleaning and set-up###
**Packages Used:** stargazer, sjPlot, sjmisc, ggplot2, lmerTest, lme4, reshape2, tidyr, doBy, psych, dplyr, nlme, summarytools, plyr, knitr, pander, DT, shiny

###View Data###
```{r, warning=F, message=F, echo=F}
datatable(df, extensions = c('FixedColumns','FixedHeader'), options = list(pageLength=10,lengthMenu = c(5, 10, 15, 20, 50, 100), dom = 'Blfrtip', scrollX = TRUE, scrollY=TRUE, fixedHeader=TRUE, fixedColumns = list(leftColumns = 2),
  initComplete = JS("
    function(settings, json) {
      $(this.api().table().header()).css({
        'background-color': '#000',
        'color': '#fff'
      });
    }")
))
```

###Description of Variables### 
####Demographic Variables####
- ID: subject ID
- Gender_1: gender (0 = male; 1 = female)
- Age_1: age
- CurrentMed_notBC_t3: current medications, excluding oral contraceptive
- CurrentMed_BC_t3: current oral contraceptive
- Smoker_0no_1yes_2yestoday_t3: smoking status (0 = no; 1 = yes; 2 = yes and smoked today)
- TodayCaff_NumCups_t3: number of caffeinated beverages today
- CurrentMenstr_t3: current menstruation (1 = yes) 
- HurricaneImpairment: continued impairment in day-to-day functioning after Hurricane Irma related to stress of hurricane experience
</br>

####Self-Reported RNT####
- RNT_1: repetitive negative thinking on the Perseverative Thinking Questionnaire at Time 1
</br>

####Gambling Task####
- muBL: choice consistency during baseline block
- lambdaBL: loss aversion during baseline block
- rhoBL: risk seeking during baseline block
- muANT: choice consistency during stress anticipation block
- lambdaANT: loss aversion during stress anticipation block
- rhoANT: risk seeking during stress anticipation block
- muREC: choice consistency during stress recovery block
- lambdaREC: loss aversion during stress recovery block
- rhoREC: risk seeking during stress recovery block

####Cortisol and Momentary Negative Affect####
- cort0-cort2: cortisol measures over course of session (nmol/L)
- na0-na2: negative affect measures over course of session (PANAS)
</br>

###Descriptive Statistics & Basic Data Visualization###
```{r, echo=F, results="asis", message=F, warning=F}

dfsm <- df[c(2:4,11:24)]

datatable(descr(dfsm), extensions = c('FixedColumns','FixedHeader'), options = list(pageLength=15, dom = 't', scrollX = TRUE, scrollY=TRUE, fixedHeader=TRUE, fixedColumns = list(leftColumns = 1),
  initComplete = JS("
    function(settings, json) {
      $(this.api().table().header()).css({
        'background-color': '#000',
        'color': '#fff'
      });
    }")
))
```

```{r,echo=F, results="asis", message=F, warning=F}
df_subset <- df2[c(5:10)]
tab <- ldply(df_subset,function(x) t(rbind(names(table(x)),table(x),paste0(prop.table(table(x))*100,"%"))))
colnames(tab) <- c("Variable","Category","Count in Category","Percent")

datatable(tab, extensions = c('FixedColumns','FixedHeader'), options = list(pageLength=15, dom = 't', scrollX = TRUE, scrollY=TRUE, fixedHeader=TRUE, fixedColumns = list(leftColumns = 3),
  initComplete = JS("
    function(settings, json) {
      $(this.api().table().header()).css({
        'background-color': '#000',
        'color': '#fff'
      });
    }")
))


```

```{r, include=F, warning=F, message=F}
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
    library(plyr)
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else length(x)
    }
    datac <- ddply(data, groupvars, .drop=.drop, .fun = function(xx, col) {
        c(N=length2(xx[[col]], na.rm=na.rm),
          mean=mean(xx[[col]], na.rm=na.rm),
          sd=sd(xx[[col]], na.rm=na.rm))
      }, measurevar)
    datac <- rename(datac, c("mean" = measurevar))
    datac$se <- datac$sd / sqrt(datac$N)
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult
    return(datac)
}
```
<br/>

###Cortisol Visualization###
```{r, warning=F, message=F}
cortvalues <- NA
cortvalues[1] <- mean(df$cort0, na.rm=T)
cortvalues[2] <- mean(df$cort1, na.rm=T)
cortvalues[3] <- mean(df$cort2, na.rm=T)

sevalues <- NA
a <- summarySE(df, measurevar="cort0", na.rm=T)
sevalues[1] <- a$se
a2 <- summarySE(df, measurevar="cort1", na.rm=T)
sevalues[2] <- a2$se
a3 <- summarySE(df, measurevar="cort2", na.rm=T)
sevalues[3] <- a3$se

Cortvalues <- NA
Cortvalues <- cbind(as.data.frame(cortvalues),as.data.frame(sevalues))

Cortvalues$Time <- c(0,1,2)
```
<br/>

```{r, warning=F, message=F, results="hide"}
p<- ggplot(Cortvalues, aes(x=Time, y=cortvalues)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=cortvalues-sevalues, ymax=cortvalues+sevalues), width=.2,
                 position=position_dodge(0.05))
```

```{r, warning=F,message=F}
# Finished line plot
p+labs(title="Average cortisol across course of experimental stress manipulation", x="Time Point", y = "Average Cortisol (nmol/L)")+ theme_classic() + scale_x_continuous(breaks=c(0,1,2), labels=c("Baseline","Stress","Recovery")) + theme(plot.title = element_text(hjust = 0.5))
```

The plot clearly indicates an average increase in cortisol during the stress induction, and a decrease in cortisol as individuals recovered from the stressor. 
<br/>


###Negative Affect Visualization###
```{r, include=F}
navalues <- NA
navalues[1] <- mean(df$na0, na.rm=T)
navalues[2] <- mean(df$na1, na.rm=T)
navalues[3] <- mean(df$na2, na.rm=T)

sevalues <- NA
a <- summarySE(df, measurevar="na0", na.rm=T)
sevalues[1] <- a$se
a2 <- summarySE(df, measurevar="na1", na.rm=T)
sevalues[2] <- a2$se
a3 <- summarySE(df, measurevar="na2", na.rm=T)
sevalues[3] <- a3$se

Navalues <- NA
Navalues <- cbind(as.data.frame(navalues),as.data.frame(sevalues))

Navalues$Time <- c(0,1,2)
```

```{r, include=F}
p<- ggplot(Navalues, aes(x=Time, y=navalues)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=navalues-sevalues, ymax=navalues+sevalues), width=.2,
                 position=position_dodge(0.05))
```

```{r, warning=F,message=F,echo=F}
p+labs(title="Average negative affect across course of experimental stress manipulation", x="Time Point", y = "Average Negative Affect (PANAS)")+ theme_classic() + scale_x_continuous(breaks=c(0,1,2), labels=c("Baseline","Stress","Recovery")) + theme(plot.title = element_text(hjust = 0.5))
```

###Gambling Task Parameter Visualization###
####Rho####
```{r, include=F}
rhovalues <- NA
rhovalues[1] <- mean(df$rho_BL_log, na.rm=T)
rhovalues[2] <- mean(df$rho_ANT_log, na.rm=T)
rhovalues[3] <- mean(df$rho_REC_log, na.rm=T)

sevalues <- NA
a <- summarySE(df, measurevar="rho_BL_log", na.rm=T)
sevalues[1] <- a$se
a2 <- summarySE(df, measurevar="rho_ANT_log", na.rm=T)
sevalues[2] <- a2$se
a3 <- summarySE(df, measurevar="rho_REC_log", na.rm=T)
sevalues[3] <- a3$se

Rhovalues <- NA
Rhovalues <- cbind(as.data.frame(rhovalues),as.data.frame(sevalues))

Rhovalues$Time <- c(0,1,2)
```

```{r, include=F}
p<- ggplot(Rhovalues, aes(x=Time, y=rhovalues)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=rhovalues-sevalues, ymax=rhovalues+sevalues), width=.2,
                 position=position_dodge(0.05))
```

```{r, warning=F,message=F,echo=F}
p+labs(title="Average rho across course of experimental stress manipulation", x="Time Point", y = "Average Rho")+ theme_classic() + scale_x_continuous(breaks=c(0,1,2), labels=c("Baseline","Stress","Recovery")) + theme(plot.title = element_text(hjust = 0.5))
```

####Lambda####
```{r, include=F}
lambdavalues <- NA
lambdavalues[1] <- mean(df$lambda_BL_log, na.rm=T)
lambdavalues[2] <- mean(df$lambda_ANT_log, na.rm=T)
lambdavalues[3] <- mean(df$lambda_REC_log, na.rm=T)

sevalues <- NA
a <- summarySE(df, measurevar="lambda_BL_log", na.rm=T)
sevalues[1] <- a$se
a2 <- summarySE(df, measurevar="lambda_ANT_log", na.rm=T)
sevalues[2] <- a2$se
a3 <- summarySE(df, measurevar="lambda_REC_log", na.rm=T)
sevalues[3] <- a3$se

Lambdavalues <- NA
Lambdavalues <- cbind(as.data.frame(lambdavalues),as.data.frame(sevalues))

Lambdavalues$Time <- c(0,1,2)
```

```{r, include=F}
p<- ggplot(Lambdavalues, aes(x=Time, y=lambdavalues)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=lambdavalues-sevalues, ymax=lambdavalues+sevalues), width=.2,
                 position=position_dodge(0.05))
```

```{r, warning=F,message=F,echo=F}
p+labs(title="Average lambda across course of experimental stress manipulation", x="Time Point", y = "Average Lambda")+ theme_classic() + scale_x_continuous(breaks=c(0,1,2), labels=c("Baseline","Stress","Recovery")) + theme(plot.title = element_text(hjust = 0.5))
```

####Mu####
```{r, include=F}
muvalues <- NA
muvalues[1] <- mean(df$mu_BL_log, na.rm=T)
muvalues[2] <- mean(df$mu_ANT_log, na.rm=T)
muvalues[3] <- mean(df$mu_REC_log, na.rm=T)

sevalues <- NA
a <- summarySE(df, measurevar="mu_BL_log", na.rm=T)
sevalues[1] <- a$se
a2 <- summarySE(df, measurevar="mu_ANT_log", na.rm=T)
sevalues[2] <- a2$se
a3 <- summarySE(df, measurevar="mu_REC_log", na.rm=T)
sevalues[3] <- a3$se

Muvalues <- NA
Muvalues <- cbind(as.data.frame(muvalues),as.data.frame(sevalues))

Muvalues$Time <- c(0,1,2)
```

```{r, include=F}
p<- ggplot(Muvalues, aes(x=Time, y=muvalues)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=muvalues-sevalues, ymax=muvalues+sevalues), width=.2,
                 position=position_dodge(0.05))
```

```{r, warning=F,message=F,echo=F}
p+labs(title="Average mu across course of experimental stress manipulation", x="Time Point", y = "Average Mu")+ theme_classic() + scale_x_continuous(breaks=c(0,1,2), labels=c("Baseline","Stress","Recovery")) + theme(plot.title = element_text(hjust = 0.5))
```


###Preliminary Analyses###
```{r}
t.test(df$cort1, df$cort0, paired=T)
t.test(df$cort2, df$cort0, paired=T)
t.test(df$cort2, df$cort1, paired=T)

t.test(df$na1, df$na0, paired=T)
t.test(df$na2, df$na0, paired=T)
t.test(df$na2, df$na1, paired=T)

t.test(df$lambda_ANT_log, df$lambda_BL_log, paired=T)
t.test(df$lambda_REC_log, df$lambda_BL_log, paired=T)
t.test(df$lambda_REC_log, df$lambda_ANT_log, paired=T)

t.test(df$rho_ANT_log, df$rho_BL_log, paired=T)
t.test(df$rho_REC_log, df$rho_BL_log, paired=T)
t.test(df$rho_REC_log, df$rho_ANT_log, paired=T)

t.test(df$mu_ANT_log, df$mu_BL_log, paired=T)
t.test(df$mu_REC_log, df$mu_BL_log, paired=T)
t.test(df$mu_REC_log, df$mu_ANT_log, paired=T)
```

###Regression Models###
####Aim 1####
```{r, warning=F, results="asis"}
hurte1 <- lm(scale(HurricaneImpairment) ~ scale(RNT_1), data=df)
stargazer(hurte1, ci=TRUE, type="html")
```

####Aim 2####
#####Baseline Models####
```{r, warning=F, results="asis"}
cortBLmod <- lm(scale(cort0) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3), data=df)
stargazer(cortBLmod, ci=TRUE, type="html")
           
naBLmod <- lm(scale(na0) ~ scale(RNT_1) , data=df)
stargazer(naBLmod, ci=TRUE, type="html")

rhoBLmod <- lm(scale(rho_BL_log) ~ scale(RNT_1) , data=df)
stargazer(rhoBLmod, ci=TRUE, type="html")

lambdaBLmod <- lm(scale(lambda_BL_log) ~ scale(RNT_1) , data=df)
stargazer(lambdaBLmod, ci=TRUE, type="html")

muBLmod <- lm(scale(mu_BL_log) ~ scale(RNT_1) , data=df)
stargazer(muBLmod, ci=TRUE, type="html")
```

####Aim 3####
####Stress Models####
```{r, warning=F, results="asis"}
cortANTmod <- lm(scale(cort1) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3), data=df)
stargazer(cortANTmod, ci=TRUE, type="html")

cortANTmod2 <- lm(scale(cort1) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3) + scale(cort0), data=df)
stargazer(cortANTmod2, ci=TRUE, type="html")

naANTmod <- lm(scale(na1) ~ scale(RNT_1) , data=df)
stargazer(naANTmod, ci=TRUE, type="html")
confint.lm(naANTmod)

naANTmod2 <- lm(scale(na1) ~ scale(RNT_1) + scale(na0), data=df)
stargazer(naANTmod2, ci=TRUE, type="html")

rhoANTmod <- lm(scale(rho_ANT_log) ~ scale(RNT_1) , data=df)
stargazer(rhoANTmod, ci=TRUE, type="html")

rhoANTmod2 <- lm(scale(rho_ANT_log) ~ scale(RNT_1) + scale(rho_BL_log), data=df)
stargazer(rhoANTmod2, ci=TRUE, type="html")

lambdaANTmod <- lm(scale(lambda_ANT_log) ~ scale(RNT_1) , data=df)
stargazer(lambdaANTmod, ci=TRUE, type="html")

lambdaANTmod2 <- lm(scale(lambda_ANT_log) ~ scale(RNT_1) + scale(lambda_BL_log), data=df)
stargazer(lambdaANTmod2, ci=TRUE, type="html")

muANTmod <- lm(scale(mu_ANT_log) ~ scale(RNT_1) , data=df)
stargazer(muANTmod, ci=TRUE, type="html")

muANTmod2 <- lm(scale(mu_ANT_log) ~ scale(RNT_1) + scale(mu_BL_log), data=df)
stargazer(muANTmod2, ci=TRUE, type="html")
```

####Aim 4####
####Recovery Models####
```{r, warning=F, results="asis"}
cortRECmod <- lm(scale(cort2) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3), data=df)
stargazer(cortRECmod, ci=TRUE, type="html")

cortRECmod2 <- lm(scale(cort2) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3) + scale(cort0), data=df)
stargazer(cortRECmod2, ci=TRUE, type="html")

cortRECmod3 <- lm(scale(cort2) ~ scale(RNT_1) +scale(Smoker_0no_1yes_2yestoday_t3) + Gender+ scale(TodayCaff_NumCups_t3) + scale(cort1), data=df)
stargazer(cortRECmod3, ci=TRUE, type="html")

naRECmod <- lm(scale(na2) ~ scale(RNT_1) , data=df)
stargazer(naRECmod, ci=TRUE, type="html")
confint.lm(naRECmod)

naRECmod2 <- lm(scale(na2) ~ scale(RNT_1) + scale(na0), data=df)
stargazer(naRECmod2, ci=TRUE, type="html")

naRECmod3 <- lm(scale(na2) ~ scale(RNT_1) + scale(na1), data=df)
stargazer(naRECmod3, ci=TRUE, type="html")

rhoRECmod <- lm(scale(rho_REC_log) ~ scale(RNT_1) , data=df)
stargazer(rhoRECmod, ci=TRUE, type="html")

rhoRECmod2 <- lm(scale(rho_REC_log) ~ scale(RNT_1) + scale(rho_BL_log), data=df)
stargazer(rhoRECmod2, ci=TRUE, type="html")

rhoRECmod3 <- lm(scale(rho_REC_log) ~ scale(RNT_1) + scale(rho_ANT_log), data=df)
stargazer(rhoRECmod3, ci=TRUE, type="html")

lambdaRECmod <- lm(scale(lambda_REC_log) ~ scale(RNT_1) , data=df)
stargazer(lambdaRECmod, ci=TRUE, type="html")

lambdaRECmod2 <- lm(scale(lambda_REC_log) ~ scale(RNT_1) + scale(lambda_BL_log), data=df)
stargazer(lambdaRECmod2, ci=TRUE, type="html")

lambdaRECmod3 <- lm(scale(lambda_REC_log) ~ scale(RNT_1) + scale(lambda_ANT_log), data=df)
stargazer(lambdaRECmod3, ci=TRUE, type="html")

muRECmod <- lm(scale(mu_REC_log) ~ scale(RNT_1) , data=df)
stargazer(muRECmod, ci=TRUE, type="html")

muRECmod2 <- lm(scale(mu_REC_log) ~ scale(RNT_1) + scale(mu_BL_log), data=df)
stargazer(muRECmod2, ci=TRUE, type="html")

muRECmod3 <- lm(scale(mu_REC_log) ~ scale(RNT_1) + scale(mu_ANT_log), data=df)
stargazer(muRECmod3, ci=TRUE, type="html")
```

